A simplified dynamic bioenergetic model for coral-\emph{Symbiodinium} symbioses and coral bleaching as an alternate stable state

### 2017-01-08 JTB_ms/JTB_ms.Rmd
Fig8 formatting

The occurrence and the magnitude of this overshoot depend on the relative inputs of carbon and nitrogen to the system -- any factor that enhances carbon-limitation (e.g. high DIN and/or low feeding) therefore magnifies the overshoot and prolongs the dysfunctional, carbon-limited state of the symbiosis (Fig. 8). On the other hand, factors that favor nitrogen-limitation, such as low external DIN, high feeding rates, will prevent an overshoot from occurring at all, since nitrogen-limitation is re-established in the system before S:H biomass ratios reach high levels. While many scenarios are possible under different environmental conditions, we illustrate the general effects of varying N and C availability on recovery from bleaching with a series of simulations that vary the N:C ratio of host's heterotrophic food source (Fig. 8): lower N:C ratios (effectively representing lower DIN and/or higher heterotrophy) favor nitrogen-limitation and more rapid recovery, while higher N:C ratios (effectively representing higher DIN and/or lower heterotrophy) favor carbon-limitation and prolonged recovery with a larger overshoot in S:H biomass.

Importantly, the re-establishment of nitrogen-limitation that characterizes the 'healthy' symbiosis is the most important diagnostic for recovery, as this is when high carbon fixation and growth rates are resumed. A high S:H biomass ratio alone does not indicate a 'healthy' state has been reached, because overall C-limitation could still persist and thus the host isn't really growing. This could explain why corals that have apparently recovered their symbiont populations after bleaching still do not grow or suffer reduced reproduction until much later (xxx). Together, these dynamics indicate that acquisition of carbon from a source other than the symbiont is extremely important for the host in order to recover from bleaching. The system can 'recover' (in terms of resuming positive growth and N-limitation) more rapidly when the coral obtains more carbon. Indeed, feeding has been shown to promote more rapid recovery from bleaching (xxx). Moreover, if the host could obtain an independent source of carbon, such as direct uptake of DOC (xxx), this could be even more important in ensuring rapid recovery from a bleached state.

>>>>

The occurrence and the magnitude of this overshoot depend on the relative inputs of carbon and nitrogen to the system -- any factor that enhances carbon-limitation (e.g. high DIN and/or low feeding) therefore magnifies the overshoot and prolongs the dysfunctional, carbon-limited state of the symbiosis (Fig. 8). On the other hand, factors that favor nitrogen-limitation, such as low external DIN and/or high feeding rates, will accelerate the re-establishment of nitrogen-limitation and prevent an overshoot from occurring at all. While many scenarios are possible under different environmental conditions, we illustrate the general effects of varying N and C availability on recovery from bleaching with a series of simulations that vary the N:C ratio of host's heterotrophic food source (Fig. 8): lower N:C ratios (effectively representing lower DIN and/or higher heterotrophy) favor nitrogen-limitation and more rapid recovery, while higher N:C ratios (effectively representing higher DIN and/or lower heterotrophy) favor carbon-limitation and prolonged recovery with a larger overshoot in S:H biomass.

Importantly, the re-establishment of nitrogen-limitation is the most important diagnostic of recovery to a 'healthy' state, as this is when high carbon fixation and growth rates are resumed. A high S:H biomass ratio alone does not indicate that a 'healthy' state has been reached, since overall C-limitation may still persist (e.g., high N:C ratios in Fig. 8). This could explain why corals that have apparently recovered their symbiont populations after bleaching still have reduced growth and reproduction rates until much later (xxx, Hughes). Therefore, acquisition of carbon from a source other than the symbiont may be extremely important for the host in order to recover from bleaching. Indeed, feeding has been shown to promote more rapid recovery from bleaching (xxx). Moreover, if the host obtains carbon from another independent source, such as direct uptake of DOC (xxx), this could be even more important in ensuring rapid recovery from a bleached state.

### 2017-01-08 JTB_ms/JTB_ms.tex
Fig8 formatting

The occurrence and the magnitude of this overshoot depend on the
relative inputs of carbon and nitrogen to the system -- any factor that
enhances carbon-limitation (e.g.~high DIN and/or low feeding) therefore
magnifies the overshoot and prolongs the dysfunctional, carbon-limited
state of the symbiosis (Fig. 8). On the other hand, factors that favor
nitrogen-limitation, such as low external DIN, high feeding rates, will
prevent an overshoot from occurring at all, since nitrogen-limitation is
re-established in the system before S:H biomass ratios reach high
levels. While many scenarios are possible under different environmental
conditions, we illustrate the general effects of varying N and C
availability on recovery from bleaching with a series of simulations
that vary the N:C ratio of host's heterotrophic food source (Fig. 8):
lower N:C ratios (effectively representing lower DIN and/or higher
heterotrophy) favor nitrogen-limitation and more rapid recovery, while
higher N:C ratios (effectively representing higher DIN and/or lower
heterotrophy) favor carbon-limitation and prolonged recovery with a
larger overshoot in S:H biomass.

Importantly, the re-establishment of nitrogen-limitation that
characterizes the `healthy' symbiosis is the most important diagnostic
for recovery, as this is when high carbon fixation and growth rates are
resumed. A high S:H biomass ratio alone does not indicate a `healthy'
state has been reached, because overall C-limitation could still persist
and thus the host isn't really growing. This could explain why corals
that have apparently recovered their symbiont populations after
bleaching still do not grow or suffer reduced reproduction until much
later (xxx). Together, these dynamics indicate that acquisition of
carbon from a source other than the symbiont is extremely important for
the host in order to recover from bleaching. The system can `recover'
(in terms of resuming positive growth and N-limitation) more rapidly
when the coral obtains more carbon. Indeed, feeding has been shown to
promote more rapid recovery from bleaching (xxx). Moreover, if the host
could obtain an independent source of carbon, such as direct uptake of
DOC (xxx), this could be even more important in ensuring rapid recovery
from a bleached state.

>>>>

The occurrence and the magnitude of this overshoot depend on the
relative inputs of carbon and nitrogen to the system -- any factor that
enhances carbon-limitation (e.g.~high DIN and/or low feeding) therefore
magnifies the overshoot and prolongs the dysfunctional, carbon-limited
state of the symbiosis (Fig. 8). On the other hand, factors that favor
nitrogen-limitation, such as low external DIN and/or high feeding rates,
will accelerate the re-establishment of nitrogen-limitation and prevent
an overshoot from occurring at all. While many scenarios are possible
under different environmental conditions, we illustrate the general
effects of varying N and C availability on recovery from bleaching with
a series of simulations that vary the N:C ratio of host's heterotrophic
food source (Fig. 8): lower N:C ratios (effectively representing lower
DIN and/or higher heterotrophy) favor nitrogen-limitation and more rapid
recovery, while higher N:C ratios (effectively representing higher DIN
and/or lower heterotrophy) favor carbon-limitation and prolonged
recovery with a larger overshoot in S:H biomass.

Importantly, the re-establishment of nitrogen-limitation is the most
important diagnostic of recovery to a `healthy' state, as this is when
high carbon fixation and growth rates are resumed. A high S:H biomass
ratio alone does not indicate that a `healthy' state has been reached,
since overall C-limitation may still persist (e.g., high N:C ratios in
Fig. 8). This could explain why corals that have apparently recovered
their symbiont populations after bleaching still have reduced growth and
reproduction rates until much later (xxx, Hughes). Therefore,
acquisition of carbon from a source other than the symbiont may be
extremely important for the host in order to recover from bleaching.
Indeed, feeding has been shown to promote more rapid recovery from
bleaching (xxx). Moreover, if the host obtains carbon from another
independent source, such as direct uptake of DOC (xxx), this could be
even more important in ensuring rapid recovery from a bleached state.
